Investigating magnetic dynamo evolution with \TESS\ field dwarfs

### 2017-10-06 ms.tex
editing first section

Of the measurable properties for a large ensemble of field stars, rotation
periods contain the most information about stellar age and provide the best
leverage for advancing our knowledge of galactic archeology as well as
exoplanet population demographics via gyrochronology.
Angular momentum is carried away though magnetically driven stellar winds,
which slows the star's rotation over cosmic time.
This rotation-based `clock' is known as gyrochronology.
Cool spots on the star's surface rotate in-to and out-of view, creating small
amplitude ($\pm$ 1\% or less) quasi-periodic changes in the stellar
brightness.
Based on results from \Kepler, using precise light curves available from the
\TESS\ mission, we expect to measure the rotation periods of around 20\% of FG
stars and 80\% of KM stars in the \tess\ Full Frame Images (FFIs)\footnote{The
difference between these percentages comes from a dramatic increase in
magnetic activity, and therefore more easily measured rotation periods,
towards lower masses.}.
For many of these stars it will be possible to infer an age from their
rotation periods via gyrochronology.
Some outstanding mysteries regarding the nature of magnetic braking were
revealed, but thus-far unanswered, by \kepler.
Firstly, a mysterious gap in the rotation period distribution of \Kepler\
dwarfs requires either a sharp transition in magnetic dynamo geometry or a gap
in the local star formation history as an explanation \citep{mcquillan2014,
davenport2017}.
Secondly, a transitioning magnetic dynamo appears to be responsible for
inefficient magnetic braking at old ages in Solar-mass stars
\citep{van-saders2016}.
To answer these questions {\bf we propose to provide a catalog of $\sim$ 1
million rotation periods for \TESS\ Full Frame Image (FFI) stars.}

>>>>

Of the measurable properties for a large ensemble of field stars, rotation
periods contain the most information about stellar age and provide the best
leverage for advancing our knowledge of galactic archeology as well as
exoplanet population demographics via gyrochronology.
Angular momentum is carried away though magnetically driven stellar winds,
which slows the star's rotation over cosmic time.
This rotation-based `clock' is known as gyrochronology.
Cool spots on the star's surface rotate in-to and out-of view, creating small
amplitude ($\pm$ 1\% or less) quasi-periodic changes in the stellar
brightness.
Thousands of photometric rotation periods have been measured from \Kepler\ and
\ktwo\ light curves, but these surveys focused on small areas of the sky.
\tess\ will open unexplored parts of the galaxy to rotational evolution
studies.
Some outstanding mysteries regarding the nature of magnetic braking were
revealed, but thus-far unanswered, by \kepler.
Firstly, a mysterious gap in the rotation period distribution of \Kepler\
dwarfs requires either a sharp transition in magnetic dynamo geometry or a gap
in the local star formation history as an explanation \citep{mcquillan2014,
davenport2017}.
Secondly, a transitioning magnetic dynamo appears to be responsible for
inefficient magnetic braking at old ages in Solar-mass stars
\citep{van-saders2016}.
To answer these questions {\bf we propose to provide a catalog of $\sim$ 1
million rotation periods for \TESS\ Full Frame Image (FFI) stars.}
